Binning Strategies and Related Loss for Large Data

### 2015-03-16 Paper.Rnw
minor edits

{\textit Rectangular binning} accomplishes this by defining a sequence of $J$ adjacent intervals, $(\beta_{j-1},\beta_{j}]$ for $j \in \{1,\dots,J\}$, which span over the range of the data. Note that half open intervals are used such that any observation falling on a bin boundary is assigned to a unique interval. Values $x_i$ exactly equal to the lowest bin boundar $\beta_0$ are grouped into the first bin to close the leftmost bound. Each observation is then mapped to a bin center, $x_j^\ast$; the midpoint for the interval to which the observation belongs.

@@

Note that this \hh{XXX which binning procedure are you referring to by 'this'?} binning procedure utilizes intervals that are open to the left and closes the outer bound of the leftmost bin. These specifications are consistent with the binning procedure used in the \texttt{hist()} function for creating histograms in base \texttt{R}. These specifications were selected for this paper, but these choices are by no means consider universal for binning.  For example, the \texttt{ggplot2} package creates histograms with intervals open to the right and does not close the outer bound of the rightmost bin.

@@

Loss of information occurs during the binning and rendering process. For the remainder of the paper we will assume that we are using colors in binned scatterplots to represent frequencies.  We distinguish two sources of loss in the construction of a binned scatterplot:
\begin{itemize}
\item {\it Spatial Loss}, $L^\text{S}$, occurs when points ($x_i$,$y_i$) \hh{XXX define $i$, $l$} in the data set are reduced to a single points at the bin centers ($x_l^\ast$,$y_l^\ast$). By displaying frequency information using tiles that are visually centered at the bin centers causes a loss of information about the exact location of the points.

>>>>

{\textit Rectangular binning} accomplishes this by defining a sequence of $J$ adjacent intervals, $(\beta_{j-1},\beta_{j}]$ for $j \in \{1,\dots,J\}$, which span over the range of the data. Note that half open intervals are used such that any observation falling on a bin boundary is assigned to a unique interval. Values $x_i$ exactly equal to the lowest bin boundary $\beta_0$ are grouped into the first bin to close the leftmost bound. Each observation is then mapped to a bin center, $x_j^\ast$; the midpoint for the interval to which the observation belongs.

@@

Note that this standard rectangular binning procedure utilizes intervals that are open to the left and closes the outer bound of the leftmost bin. These specifications are consistent with the binning procedure used in the \texttt{hist()} function for creating histograms in base \texttt{R}. These specifications were selected for this paper, but these choices are by no means consider universal for binning.  For example, the \texttt{ggplot2} package creates histograms with intervals open to the right and does not close the outer bound of the rightmost bin.

@@

Loss of information occurs during the binning and rendering process. For the remainder of the paper we will assume that we are using colors in binned scatterplots to represent frequencies.  We distinguish two sources of loss in the construction of a binned scatterplot:
\begin{itemize}
\item {\it Spatial Loss}, $L^\text{S}$, occurs when points ($x_i$,$y_i$) \hh{XXX define $i$, $l$} in the data set are reduced to a single points at the bin centers ($x_\ell^\ast$,$y_\ell^\ast$). By displaying frequency information using tiles that are visually centered at the bin centers causes a loss of information about the exact location of the points.
